ms_results

### 2023-01-06 ms_figures.Rmd
update trade flow trend section and figure

region_trend_matrix_method <- artis_region_method %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
  geom_area() +
  scale_fill_manual(values = habitat_source_colors) +
  labs(x = "", y = "Trade (million t, live weight)", color = "Ave. annual change in export \n(1000 t live weight)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        strip.text.y = element_blank()) +
  facet_grid(rows = vars(exporter_region), cols = vars(importer_region), 
             labeller = label_wrap_gen(width = 5), scales = "free")

# Create basic plot for the legend
artis_region_method %>%
  group_by(year, exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ggplot(aes(x = year, y = live_weight_t, fill = exporter_region)) +
  geom_area() +
  scale_fill_manual(values = region_colors) +
  labs(fill = "Region") +
  theme(legend.position = "bottom")
```

@@

artis_region <- artis_region %>%
  left_join(region_slopes, by = "exporter_importer")

@@

```{r region_trend_lollipop, fig.height=4, fig.width=8}
region_slopes_ranked <- region_slopes %>%
  mutate(exporter_importer = fct_reorder(exporter_importer, slope)) %>%
  ggplot(aes(x = slope/1000, y = exporter_importer, color = slope, fill = slope)) +
  geom_segment(aes(xend = 0, yend = exporter_importer)) +
  geom_point() +
  scale_color_gradient2(low = "#E24027", mid = "grey70", high = "#1B708F") +
  scale_fill_gradient2(low = "#E24027", mid = "grey70", high = "#1B708F") +
  labs(y = "", x = "Ave. annual change in export (1000 t live weight)") +
  theme_bw() +
  theme(legend.position = "blank")

ggarrange(region_trend_matrix_method, region_slopes_ranked, widths = c(1.5,1))
```

```{r, fig.height=3, fig.width=4}
exports_by_region <- artis_region %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = importer_region)) +
  geom_area() +
  labs(x = "", y = "", fill = "Importer") +
  scale_fill_manual(values = region_colors) +
  facet_wrap(~exporter_region, ncol = 1) +
  theme_bw() +
    theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

imports_by_region <- artis_region %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = exporter_region)) +
  geom_area() +
  labs(x = "", y = "", fill = "Exporter") +
  scale_fill_manual(values = region_colors) +
  facet_wrap(~importer_region, nrow = 1) +
  theme_bw() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.text.x = element_blank())

png("outputs/fig2a.png")
ggarrange(ggarrange(imports_by_region, NULL,
                    ncol = 2, nrow = 1,
                    widths = c(4, 1)),
          ggarrange(region_trend_matrix_method, exports_by_region, 
                    ncol = 2, nrow = 1,
                    widths = c(4, 1)),
          ncol = 1, nrow = 2, 
          heights = c(1, 4))
dev.off()
```

>>>>

# Create basic plot for the legend
# artis_region_method %>%
#   group_by(year, exporter_region) %>%
#   summarise(live_weight_t = sum(live_weight_t)) %>%
#   ggplot(aes(x = year, y = live_weight_t, fill = exporter_region)) +
#   geom_area() +
#   scale_fill_manual(values = region_colors) +
#   labs(fill = "Region") +
#   theme(legend.position = "bottom")
```

@@

artis_region <- artis_region %>%
  left_join(region_slopes, by = "exporter_importer") %>% 
  mutate(intraregion = case_when(
           exporter_region == importer_region ~ "intraregional trade",
           exporter_region != importer_region ~ "interegional trade"))

@@

# SI Fig X: Bilateral region trends lollypop plot

```{r region_trend_lollipop, fig.height=4, fig.width=3}
region_slopes_ranked <- artis_region %>%
  select(-year, -live_weight_t) %>%
  distinct() %>%
  mutate(exporter_importer = fct_reorder(exporter_importer, slope)) %>%
  ggplot(aes(x = slope/1000, y = exporter_importer, color = intraregion)) +
  geom_segment(aes(x=0, xend=slope/1000, y=exporter_importer, yend=exporter_importer, color = intraregion)) +
  geom_point(size=3) +
  scale_y_reordered() +
  scale_color_manual(values = habitat_source_colors[c(1,4)]) +
  labs(y = "", x = "Ave. annual change in export (1000 t live weight)", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")

region_slopes_ranked

```{r, fig.height=3, fig.width=5}
region_trend_matrix_method <- artis_region_method %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
  geom_area() +
  scale_fill_manual(values = habitat_source_colors) +
  labs(x = "", y = "Trade (million t, live weight)", color = "Ave. annual change in export \n(1000 t live weight)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        legend.position = "none",
        strip.background = element_blank(),
        plot.margin=unit(c(0,0,0,0), "cm")) +
  facet_grid(rows = vars(exporter_region), cols = vars(importer_region), 
             labeller = label_wrap_gen(width = 5), scales = "free")

exports_by_region <- artis_region %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = importer_region)) +
  geom_area() +
  labs(x = "", y = "", fill = "Importer") +
  scale_fill_manual(values = region_colors) +
  facet_wrap(~exporter_region, ncol = 1) +
  theme_bw() +
    theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.margin=unit(c(0,0.5,0,0), "cm"))

imports_by_region <- artis_region %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = exporter_region)) +
  geom_area() +
  labs(x = "", y = "", fill = "Exporter") +
  scale_fill_manual(values = region_colors) +
  facet_wrap(~importer_region, nrow = 1) +
  theme_bw() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.text.x = element_blank(),
        plot.margin=unit(c(0.5,0,0,0), "cm"))

png("outputs/fig2a.png", width = 6.5, height = 5, res = 500, units = "in")
ggarrange(ggarrange(imports_by_region, NULL,
                    ncol = 2, nrow = 1,
                    widths = c(4.5, 1.25), labels = c("a", "")),
          ggarrange(region_trend_matrix_method, 
                    ggarrange(NULL, exports_by_region,
                              ncol = 1, heights = c(0.3, 4)), 
                    ncol = 2, nrow = 1,
                    widths = c(4.5, 1), labels = c("b", "c")),
          ncol = 1, nrow = 2, 
          heights = c(1, 4.5))
dev.off()
```

### 2023-01-06 ms_results.Rmd
update trade flow trend section and figure

```{r covid_effect}
artis %>% 
  filter(year %in% 2019:2020) %>% 
  group_by(year, habitat_method) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>% 
  pivot_wider(names_from = year, values_from = live_weight_t, names_prefix = "year") %>% 
  mutate(change = (year2019-year2020)/year2020) %>% 
  arrange(desc(change))
```

@@

# Regionalization trends
# Changes in trade flow volumes
bilateral_habitat_method_summary %>% 
  filter(exporter_region != "Other nei", importer_region != "Other nei",
         year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, exporter_region, importer_region) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
  mutate(change = end-beginning,
         from_to = paste(exporter_region, " to ", importer_region), 
         intraregion = case_when(
           exporter_region == importer_region ~ "intraregional trade",
           exporter_region != importer_region ~ "interegional trade"
         )) %>% 
  arrange(desc(change)) %>% 
  ungroup() %>%
  ggplot(aes(x = change/1000000, y = reorder(from_to, change), color = intraregion)) +
  geom_segment(aes(x=0, xend=change/1000000, y=reorder(from_to, change), yend=reorder(from_to, change), color = intraregion)) +
  geom_point(size=3) +
  scale_y_reordered() +
  scale_color_manual(values = habitat_source_colors[c(1,4)]) +
  labs(y = "", x = "Live weight (million t)", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")

# Changes in trade flow percents
bilateral_habitat_method_summary %>% 
  filter(exporter_region != "Other nei", importer_region != "Other nei",
         year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, exporter_region, importer_region) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
  mutate(change = 100*(end-beginning)/beginning,
         from_to = paste(exporter_region, " to ", importer_region), 
         intraregion = case_when(
           exporter_region == importer_region ~ "intraregional trade",
           exporter_region != importer_region ~ "interegional trade"
         )) %>% 
  arrange(desc(change)) %>% 
  ungroup() %>%
  ggplot(aes(x = change, y = reorder(from_to, change), color = intraregion)) +
  geom_segment(aes(x=0, xend=change, y=reorder(from_to, change), yend=reorder(from_to, change), color = intraregion)) +
  geom_point(size=3) +
  scale_y_reordered() +
  scale_color_manual(values = habitat_source_colors[c(1,4)]) +
  labs(y = "", x = "Percent change", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")

@@

Since 1996, nearly all regional trade pairs increased, other than South America. The largest increases occurred for X, Y and Z. At the country level, X% of trade pairs increased, with the most increases for marine/inland aquaculture/capture. The declining exports from South America mirror declining marine catch over this period. [Suggestions for discussion points and context related to this?].

>>>>

```{r covid_effect}
artis %>% 
  filter(year %in% 2019:2020) %>% 
  group_by(year, habitat_method) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>% 
  pivot_wider(names_from = year, values_from = live_weight_t, names_prefix = "year", values_fill = 0) %>% 
  mutate(change = (year2019-year2020)/year2020) %>% 
  arrange(desc(change))
```

@@

# Regionalization trends
# Changes in trade flow volumes
bilateral_habitat_method_summary %>% 
  filter(exporter_region != "Other nei", importer_region != "Other nei",
         year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, exporter_region, importer_region) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t, values_fill = 0) %>% 
  mutate(change = end-beginning,
         from_to = paste(exporter_region, " to ", importer_region), 
         intraregion = case_when(
           exporter_region == importer_region ~ "intraregional trade",
           exporter_region != importer_region ~ "interegional trade"
         )) %>% 
  arrange(desc(change)) %>% 
  ungroup() %>%
  ggplot(aes(x = change/1000000, y = reorder(from_to, change), color = intraregion)) +
  geom_segment(aes(x=0, xend=change/1000000, y=reorder(from_to, change), yend=reorder(from_to, change), color = intraregion)) +
  geom_point(size=3) +
  scale_y_reordered() +
  scale_color_manual(values = habitat_source_colors[c(1,4)]) +
  labs(y = "", x = "Live weight (million t)", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")

# Calculate percent of pairs with increased trade flows
bilateral_flow_change <- bilateral_habitat_method_summary %>% 
  filter(exporter_region != "Other nei", importer_region != "Other nei",
         year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, exporter_iso3c, importer_iso3c) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t, values_fill = 0) %>% 
  mutate(change = end-beginning) %>%
  arrange(desc(change))

percent_flow_increases <- 100*(nrow(bilateral_flow_change %>% filter(change > 0)))/nrow(bilateral_flow_change)

# Calculate percent of pairs with increased trade flows by production method and habitat
bilateral_flow_change_method <- bilateral_habitat_method_summary %>% 
  filter(exporter_region != "Other nei", importer_region != "Other nei",
         year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, exporter_iso3c, importer_iso3c, habitat_method) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t, values_fill = 0) %>% 
  mutate(change = end-beginning) %>%
  arrange(desc(change))

percent_flow_increases_marine_capture <- 100*(nrow(bilateral_flow_change_method %>% 
                                                     filter(habitat_method == "marine capture",
                                                            change > 0)))/nrow(bilateral_flow_change)

@@

Since 1996, nearly all regional trade pairs increased, other than South America to Europe and Asia and within North America. The largest average annual growth increases occurred for trade within Asia, within Europe and between Europe and Asia. At the country level, `r round(percent_flow_increases, 1)`% of trade pairs increased, with the most increases for marine/inland aquaculture/capture. The declining exports from South America mirror declining marine catch over this period. [Suggestions for discussion points and context related to this?].
